---
title: "Introducing the hellno package"
author: "Peter Meißner"
date: "2015-12-14"
output: 
  html_document: 
    keep_md: yes
---

[![Travis-CI Build Status](https://travis-ci.org/petermeissner/hellno.svg?branch=master)](https://travis-ci.org/petermeissner/hellno)

# Introduction

Base R's once done choice of setting stringsAsFactors within
  `data.frame()` and `as.data.frame()` to TRUE by default is a design decision
  that makes sense (more efficient storage, building statistical models with 
  factors makes sense) on the one hand and on the other hand is supposedly the 
  most often complained about piece of code in the R infrastructure. 
  A search through the source code of all CRAN packages in December 2015 
  `https://github.com/search?utf8=%E2%9C%93&q=user%3Acran+stringsAsFactors&type=Code` 
  resulted in 3,795 results for mentions of `stringsAsFactors` and most of 
  them simply set the value to `FALSE`.
  The hellno package provides an explicit solution to the problem without 
  changing R itself or having to mess around with options. It tries to solve 
  this problem by providing alternative `data.frame()` and `as.data.frame()` 
  functions that are in fact simple wrappers around base R's `data.frame()` and 
  `as.data.frame()` with `stringAsFactors` option set to `HELLNO` 
  (equals to `FALSE`) by default.

# Using hellno interactively

**R's default behaviour...**
```{r}
df1 <- data.frame(a=letters[1:3])
df1$a
class(df1$a)
```


**R's default behaviour after loading the package**
```{r}
library(hellno)

df2 <- data.frame(a=letters[1:3])
df2$a
class(df2$a)
```

# Using hellno for package development

While using the hellno package in interactive R mode is nice, in fact it could 
have been achieved simply by doing something like this: `options("stringsAsFactors"=FALSE)`. 
The strength of hellno is that it can be imported when writing packages and therefore 
providing `as.data.frame()` and `data.frame()` with `stringsAsFactors` Option 
consistently set to `FALSE`. Once imported `stringsAsFactors=FALSE` will be the default for all uses of `data.frame()` and `as.data.frame()` within all package functions BUT NOT OUTSIDE OF IT. 
Thus it provides a way to ease programming while also ensuring that package users might 
still choose which flavor of `stringsAsFactors` they like best. 


Let us see how this works following a little example. Again, let us start with loading hellno package:

```{r} 
library(hellno)
data.frame(a=letters[1:2])$a 
```

As shown before, character vector are not transformed to factor.

We unload hellno again to start clean. 

```{r}
unloadNamespace("hellno")
```

Now we install the hellnotest package from Github and load it. The package uses hellno internally in two functions. While internal uses of `data.frame()` and `as.data.frame()` will work with `stringsAsFactors=FALSE` as default this does not change how things work everywhere else. 


```{r}
if( !("hellnotests" %in% installed.packages()) ){
  devtools::install_github("petermeissner/hellnotests")
}

library(hellnotests)
data.frame(a=letters[1:2])$a 
```

While all functions within the package use hellno's alternative implementations:

```{r}
hellno_df
```

... and hence for them string conversion is no matter anymore:

```{r}
hellno_df()
```


... and once again to bring the point home: 

```{r}
data.frame(a=letters[1:2])$a 
```

***WRITING PACAKGES WITH HELLNO DOES NOT CHANGE OUTSIDE BEHAVIOR.***

