---
title: "README"
author: "Peter Meißner"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction

Base R's once done choice of setting stringsAsFactors within
  `data.frame()` and `as.data.frame()` to TRUE by default is a design decision
  that makes sense on the one hand and on the other hand is supposedly the most
  often complained about piece of code in the R infrastructure. The hellno
  package provides an explicit solution to the problem without changing R itself 
  or having to mess around with options. It tries to solve this problem by 
  providing alternative `data.frame()` and `as.data.frame()` functions that are 
  in fact simple wrappers around base R's `data.frame()` and `as.data.frame()` 
  with `stringAsFactors` option set to `HELLNO` (equals to `FALSE`) by default.

# Using the package

**R's default behaviour...**
```{r}
df1 <- data.frame(a=letters[1:3])
df1$a
class(df1$a)
```


**R's default behaviour after loading he package**
```{r}
library(hellno)

df2 <- data.frame(a=letters[1:3])
df2$a
class(df2$a)
```

# Programming with hellno

While using the hellno package in interactive R mode is nice, in fact it could 
have been achieved simply by doing something like this: `options("stringsAsFactors"=FALSE)`. 
The strength of hellno is that it can be imported when writing packages and therefore 
providing `as.data.frame()` and `data.frame()` with `stringsAsFactors` Option 
consistently set to `FALSE` within the package and its functions BUT NOT OUTSIDE OF IT. 
Thus it providdes a way to ease programming while also ensuring that package users might 
still choose which flavor of `stringsAsFactors` they like best. 





